---
title: "DS5220 Project 1 "
author: "Chase Kenyon"
date: "02/08/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

R has excellent visualization capabilities, especially with the `ggplot2` package.  Please read Chapter 3 of  [*R for Data Science [GW],* Garrett Grolemund, Hadley Wickham](https://r4ds.had.co.nz), and complete the exercises below after you finish each section.  Edit the markdown file which came with this html directly.  Make sure to enter your R code in the chunks following each question to demonstrate your answers.  Follow each code block with a text description of your solution.  Answers without demonstration will be given little credit.  Code with no description (if requested) will be given little credit.

## 3.2.4 Exercises

1. Run ggplot(data = mpg). What do you see?  

```{r problem 3.2.4.1}
library(tidyverse)
ggplot(data = mpg)
```
At this point I only see a gray box.

2. How many rows are in mpg? How many columns? Demonstrate how you obtained your answers using R.

```{r problem 3.2.4.2}
mpg

```
Rows: 234
Cols: 11
I called MPG to display a tibble of MPG, which states the dimensions above the data.


3. What does the `drv` variable describe? Read the Help Panel in RStudio  by typing `?mpg` in the Console Panel to find out.  (You will see no output from RMarkdown here.)  Produce a description of `drv` by typing `mpg` below.

```{r problem 3.2.4.3}
mpg
```
drv stands for the drive train of the car model, where f = front wheel drive, r = rear wheel drive, anddd 4 = four wheel drive.



4. Make a scatterplot of `hwy` vs `cyl` using `geom_point`.

```{r problem 3.2.4.4}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```

5. What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?

```{r problem 3.2.4.5}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```
Since both variables are categorical, you simply get a plot showing which classes have which associated drive trains, but not any data about frequency or other potentially useful metrics.

## 3.3.1 Exercises

1. Fix the code in problem 3.3.1.1, and enter it below

```{r problem 3.3.1.1}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```
We needed to move the right parenthesis to close the aes section, putting color inside aes allows you to map it to another parameter of your dataset, but to set it all to one color, it should be a function of geom_point

2. Which variables in `mpg` are categorical? Which variables are continuous? (Hint: type `?mpg` in the Console Panel to read the documentation for the dataset in the Help Panel). How can you see this information when you run `mpg`?

```{r problem 3.3.1.2}
mpg
```
manufacturer, model, trans, drv, fl, and class are categorical, which you can tell by the <chr> label in the tibble displayed by 'mpg'

3. Using your R code for 3.3.1.1, map a continuous variable to `color`, `size`, and `shape`. How does the aesthetic `shape` behave differently for mappings to `fl` and `displ`?

```{r problem 3.3.1.3a}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = hwy, size = hwy, shape = fl))

#ggplot(data = mpg) + 
#  geom_point(mapping = aes(x = displ, y = hwy, color = hwy, size = hwy, shape = displ))
```
Shape does not work with a continuous variable mapped to it, so fl works, while displ does not.

Mapping `displ` to `color` and `size` produce meaningful results. Mapping `displ` to `shape` produces a warning. Assigning the aesthetic `shape` to `fl` produces the following.

```{r problem 3.3.1.3b}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = fl))
```
Different shapes of data points for the different categories of fuel type.

4. What happens if you map the same variable to multiple aesthetics? Use an example from your answers to problem 3.3.1.3.
```{r problem 3.3.1.3c}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = hwy, size = hwy))
```
This works and will map both aesthetics to the same variable, which can create some redundancy, but may also be easier to visualize in some cases.

5. What does the `stroke` aesthetic do? What shapes does it work with? (Hint: use`?geom_point`) Try it with `shape=21` and `stroke=displ` in your  code from 3.3.1.1.

```{r problem 3.3.1.5}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, stroke=displ), color = "blue")

```
Shape = 21 gave me continuous variable error, but stroke seems to control the size of the 'pen' used to draw the point, not dissimilar to the 'size' aesthetic.

6.What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?  Try this by modifying your code in problem 3.3.1.1.

```{r problem 3.3.1.6}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```
This creates a condition where displ < 5 is one color, and displ >= 5 is another color, based on their T/F conditions.

